\section{The institutional setting of legislative debate in the Cámara} [ca 1500 words]

An overview of the structure of legislative debate shows members who have abdicated most formal speech rights to their parties. The Cámara's Rules \citep{reglamentoDipMx.2019} set most prescriptions for debate, with general guidelines in Congress' Organic Law \citep{loceum.2019}.

%% Who controls the legislative agenda and allocates speaking time?
%% 1. Who can allocate speaking time to individual MPs? Who controls the legislative agenda – government, parties, MPs, some legislative structure specifically in charge with agenda-setting? In answering this question, you should aim to discuss both formal rules (constitutional rules, rules of procedure), party internal rules (if they exist), and informal rules.
%% - party leaders (junta selects), member (self-selects), president (duration), in that order of importance

\subsection{The boards}
There are two key actors in the legislative process, the Junta and the Mesa. The *Junta de Coordinación Política* is the Cámara's top decicion-making organ. The leaders of all parties with no fewer than five deputies are represented. The majority leader presides the Junta throughout the term. In the absence of a majority party, however, the leaders of the top-three seatholding parties preside the Junta, alternating one year each. The Junta appoints and replaces committee members, prepares each session's order of the day (/orden del día/), and in general reaches and enforces party leader agreements. It decides by majority rule, with members' votes weighted relative to group sizes in the plenary. So majority status is crucial to control the Junta \citep[cf.][]{cox.mccubbins.2005}.

The *Mesa Directiva* is the chamber's steering board. The Mesa chair is the Cámara president ex-officio. The Mesa makeup has consensual traits, regardless of there being a majority party or not. It is elected yearly by two-thirds supermajority of Cámara members from candidates proposed by the Junta. While Mesa members can reelect, the chair must rotate between the top-three seatholding parties, one year each. 

%The president recognizes speechmakers and presides over Cámara debates. The Mesa includes deputy presidents and at least one secretary from each parliamentary group. The Mesa follows the session's agenda in the Day's Order (/Orden del Día/) that is mostly set by the Junta. 

% I use the terms party and parliamentary group interchangeably, but groups are in fact more restrictive. A group (/grupo parlamentario/) needs no fewer than five deputies to earn and retain Junta representation. Moreover, groups can only form at the term's outset. Members can freely leave one group and may join another existing group at any time, with immediate effects in vote weights. But defectors cannot form new groups---no doubt raising the cost of lone turncoats. 

Agenda control is frail. First, every committee report is guaranteed floor consideration and must be included in the order. If committees were adequate agents of the Junta majority, they might serve as gatekeepers by denying unwanted bills a report. But the Junta is required to distribute committee chairs (as well as committee seats) proportionally among the parties, so some committees are bound to be preference outliers.

Second, the open rule is the default for bill consideration in the floor. Debate takes place in two stages. The entire bill is first examined /en lo general/, then articles are considered individually for amendment or deletion /en lo particular/ \citep[see][]{heller.weldon.nd}. Members can always reserve articles for deletion or amendment, denying the Junta a useful prededural tool common in other assemblies: the closed rule \citep[eg.,][]{cox.2006,weingast.1992,magar-palanza-sin-Pdt-fast-track-chile-2021jop}.

%% Do individual members have debate rights without party leadership permission?
%% 2. Do individual MPs have a guaranteed right to participate in a debate without party leadership permission? If yes, how much time and in which type of debate? 
%% - yes, within limits 
%% 4. Do particular subsets of MPs (e.g. spokesperson or committee members) have formal rights that entitle them to take the floor?
%% - yes, committee members and leader-appointed party speakers

Third, and most relevant, speakers can self-select. Individual members are entitled to take the floor when recognized by the presiding officer, for a duration set by rules or by party agreements. Party leaders allocate speaking time to a list of speakers but cannot preclude others from adding their names to that list, making debate resemble first-come-first-serve once parties have spoken. 

%%Formally, deputies enjoy equal rights and duties. With respect to deliberation, all are entitled to take the floor when recognized by the presiding officer, for a duration set by rules or by party agreements. However, as in assemblies worldwide, members with status tend to have many more rights, the rest many more duties. What defines status changes from one assembly to the next \citep{cox.2006}. In the Cámara, party leaders in general, and majority party leaders in particular sit atop the status pyramid. But when it comes to speech, individual members retain debate rights that set the basis for minority rights. 

%% 5. What are the existing types of debates? Please select the 5 most important types of debates and produce a table where you cover the name of the debate (in English and the native language of the country), its goals, particular organization rules to that kind of debate, and the total number of minutes that rules assign to each debate.

\subsection{The structure of debate}
Rules set limits for different kinds of debate summarized in the Table. The first entry refers drafters of new legislation, who who get first recognition to take the floor in order to persuade fellow lawmakers. The time limit is ten minutes when the draft is a new law, five minutes when it amends existing statutes. Deputies who wish to debate then get five minutes each. Bills that cannot be presented before the session ends migrate to the next day's order upon author's request /viva voce/ (otherwise they are referred to committee.) The rightmost columns report who selects the speaker---self-selection by drafting a bill, in this case---and who, if anyone, can veto the speaker's recognition---no one here. 

\begin{table}
|----------------------------------------+---------------+---------+------------+------------|
| Debate type (in Spanish)               | Goal          | Durat.  | Selector   | Veto       |
|----------------------------------------+---------------+---------+------------+------------|
| 1. Introduce legislation (iniciativa)  | Author        |         |            |            |
| - a new law                            | presents      | - 10'   | - member   | - no       |
| - amend a law                          | the bill      | - 5'    | - member   | - no       |
|----------------------------------------+---------------+---------+------------+------------|
| 2. Committee report (dictamen)         | Move          |         |            |            |
| - Debate en lo general vs SQ, chair    | for floor     | - 10'   | - comm.maj | - pres.^1  |
| -   "             "      "   others    | consideration | - 5'    | - members  | - pres.^1  |
| - Amendments (debate en lo particular) |               | - 5'    | - members  | - no       |
| - negative report                      |               | - 3'    | - comm.maj | - pres.^1  |
|----------------------------------------+---------------+---------+------------+------------|
| 3. Resolutions (puntos de acuerdo)     | Position      |         |            |            |
| - standard, author                     | taking        | - 10'   | - member   | - comm.maj |
| - urgent, author (obvia resolución)    |               | - 5'    | - Junta    | - floor    |
| - other speakers                       |               | - 3'    | - party    | - no       |
|----------------------------------------+---------------+---------+------------+------------|
| 4. Current events (agenda política)    | Position      | < 2hrs  |            |            |
| - Junta proponent                      | taking        | - 10'   | - Junta    | - no       |
| - other speakers                       |               | - 5'    | - member   | - no       |
|----------------------------------------+---------------+---------+------------+------------|
| ^1 = President can delay/prevent speech by granting recommit.                                |
|--------------------------------------------------------------------------------------------|
\caption{Types of debate}
\end{table}

Other speech types grant right of first recognition differently. Debate /en lo general/ grants it to the reporting committee chairperson or designated handler of the report for ten minutes (fifteen in constitutional amendments). The Cámara president can delay debate by recommiting the bill---and possibly prevent it if the committee kills the bill. /En lo particular/ amendments and Cámara resolutions grant it to the proposing member. 

Party-appointed speakers get five minutes each, in reverse-size order, after the first /en lo general/ speech. Then members who request it then get five minutes each, the president arranging them in rounds, one for one against. After six such rounds, the floor can either proceed to vote, or continue with blocks of three such rounds. When the report deals with issues of great interest, debate can go on for several hours.

%Floor (108) must approve member requests to consider an article separately from general discussion (deletion amendment?)

%Votos en lo particular = another window for dissenters. Minority report. Minority reports get floor consideration only when floor rejects the committee report. 

%Discusión en lo particular = introduce amendments to committee report (who can do it? who can block? 110 silence suggests that any member can do it --- open rule unless rules are suspended?) Proposer gets 5 minute speech, then for/againsts get 5 minute each.

Cámara resolutions (/proposiciones con punto de acuerdo/) are taylor-made for members' position-taking needs, conditional on party leader support. If adopted, resolutions become the opinion of the chamber on some specific issue. But they require urgent status (/urgente u obvia resolución/) in order to avoid committee referral and move directly to the floor; only the Junta can request that the floor grants urgent status to at most two resolutions per session. If granted, the proposer takes the floor for five minutes. Parties then appoint one speaker each, for three minutes. The floor can then decide to vote, or open a rounds of debate with self-appointed speakers.

Current events (/agenda política/) are party leaders' position-taking venue. The Junta determines up to two themes for debate before consideration of reports and bills, party leaders appointing one speaker each. The promoting party speaker gets first recognition for 10 minutes, others 5 minutes each, and talk in reverse-size order. Current events debate cannot exceed two hours per session. 

%3. How is the debate structured? What are the rules of engagement? Do MPs engage in back-and-forth talk? By contrast, do MPs deliver a debate without interruptions?
%- Back-and-forth talk is very rare, rules make for a sequence of short speeches. Once recognized, a member can speak for a limited time (extendable with President's assent). The speechmaker decides if interrupting her is kosher. 

\subsection{Recognition-granting motions}
Debate under such rules becomes a succession of punctuated, mostly uninterrupted short speeches. Members can approximate back-and-forth talk, at least occasionally, by catching the president's eye from their seats in order to interrupt with a motion. The president has discretion to deny, or grant up to three minutes to elaborate. Such motions are distinct from points of order (which members can also make, see Reglamento art. 114 for typified motions). They grant recognition to speak. One (/cuestionamiento al orador/) to interrogate the speaker, who must also accept the question be made. Another is (/alusiones personales/), to give right of reply to alluded members by recognizing them right after the speaker ends. And (/rectificación de hechos/) wind up an additional name at the end of the list of speakers. 

\subsection{Party discipline as alternative to centralized agenda power}
The Cámara's debate rules are ill-designed to prevent plenary bottlenecks \citep{cox.2006}. Even in the presence of a majority party, individual members retain speaking rights that water down attempts by the Junta to cartelize the legislative process. So how does the Cámara prevent dilatory motions to get things done? The answer is parties. Party discipline operates as an alternative to agenda cartelization in many systems \citep{prata.2006}. 

Cohesion is near perfect across parties. \citet{tellez-del-rio.2018} computed frequencies with which deputies voted against a majority of their party (excluding unanimous votes). The mean he reports for the 1997--2018 period stands at 2 percent, 3.4 percent when abstentions are coded as votes against the party majority (p. 25). 

% \citep{weldon.2005,rosas.langston.2011,cantuDesposatoMagar.MxRcv.2014}

Three former deputies from the larger parties offered quick impressions on internal party speech rules upon request. One commonality (at least in this very small sample) is the informal erosion of formal individual members' debate rights in favor of centralized speech allocation \citep[cf.][]{cox.1987}. The PAN relies on a debate whip (subcoordinador de debate parlamentario) in charge of selecting speakers in debates. When two members wish to speak at once, the whip would let them figure who would get the party's slot in the debate, who would then speak for or against.\footnote{Email exchange with Fernando Rodríguez Doval, June 17th, 2020.} The PRI leadership sets apart issues of party interest, appointing every speaker when debated. Members would communicate their wish to speak on unwhipped issues to their state caucus leader, who would seek authorization with party whips.\footnote{Email exchange with Lupita Vargas Vargas, June 17th, 2020.} Rules give parties one speaking slot each in many debates, regardless of size. Distributive conflict over speech is therefore more acute for larger parties, with longer speaker lists. A must for a member dissenting from ``party mandarins'' is a solid knowledge of the Rules. That member can thus make individual speaking rights effective by introducing suspensive motions or amendments, both of which come equipped with recognition to take the floor./footnote{Email exchange with a former deputy from the left, who answered on condition of anonimity, June 17th, 2020.} 

Party leaders move the strings of lawmaking. Their influence, however, derives almost exclusively from party discipline (near-perfect across the board) and not from agenda power (which is quite diffuse). 

\subsection{Minority rights}
%% To conclude this section, please classify the country along Proksch and Slapin’s (2015) contribution. The authors suggest that there are two extreme poles, depending on rules of procedure and electoral system organization. As described by Proksch and Slapin (2015: 96), in some systems, individual MPs are ‘guaranteed access to speaking time’, and ‘backbenchers are granted equal time as party leaders’. In other systems, the rules severely restrict individuals’ access to the floor and parties draft speakers lists, which gives party leaders much more control on this question. A number of countries fall, however, somewhere between these two extreme categories, giving some opportunity for individual MPs to access the floor, but favoring party lists. Please classify your country accordingly. If your country was included in Proksch and Slapin’s study, please refer to this classification when discussing your country’s institutional setting.
Arg here: While parties sit atop the status pyramid, their agenda power is relatively limited. 
Two problems: can't easily prevent unwanted speeches as members retain right to talk; can't easily remove unwanted motions as committee reports proceed automatically to floor discussion. So negative agenda power is diffuse. 
Parties in fact rely on discipline towards leadership to avoid plenary bottlenecks and get legislation done. 

suspension of rules typified only for discharge, two-thirds

The constitution sets the quorum at half chamber membership.

Reglamento amendments by 2/3 vote

Suspension of rules by Conferencia always a choice, but only typified for committee of the whole. Art 77 cpeum. Risks toma de tribuna.

Presiding officer can summon police to restore order. 
Can summon public force, but in practice never used. 
Can kill the mike, but others can raise their voices

4. Para atender una situación no prevista en el Reglamento, el Presidente podrá dictar una resolución de carácter general, siempre que haya la opinión favorable de la Mesa Directiva y de la Junta. En caso contrario, este tipo de resoluciones sólo tendrán efecto con la aprobación de la mayoría simple del Pleno.




